--- 
title: "PSYC4780 Course Notes"
author: "David J. Stanley"
date: "`r Sys.Date()`"
documentclass: krantz
bibliography:
- book.bib
- packages.bib
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
site: bookdown::bookdown_site
description: Course notes for PSYC 6060
github-repo: dstanley4/psyc6060bookdown
graphics: yes
---

```{r setup, include=FALSE}
options(
  htmltools.dir.version = FALSE, formatR.indent = 2, width = 55, digits = 4
)

# install the packages needed by this book; you fill out c(), e.g. c('ggplot2', 'dplyr')
lapply(c('xfun'), function(pkg) {
  if (system.file(package = pkg) == '') install.packages(pkg)
})
```

# Preface {-}

These notes are a resource for students in my PSYC*4780 class. The notes are a subset of material that I am creating for a public domain statistics book to be released sometime in the distant future.

## Software information and conventions {-}

I used the **knitr**\index{knitr} package [@xie2015] and the **bookdown**\index{bookdown} package [@R-bookdown] to compile my book. My R session information is shown below:

```{r}
xfun::session_info()
```

Package names are in bold text (e.g., **rmarkdown**), and inline code and filenames are formatted in a typewriter font (e.g., `knitr::knit('foo.Rmd')`). Function names are followed by parentheses (e.g., `bookdown::render_book()`).
